Study of Intra- and Inter-Job Interference on Torus Networks

### 2016-07-13 app_study.tex
final pass

In this work, we select three applications from the DOE Design Forward Project. 
Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. 
We believe that the communication patterns of these applications 
are representative of a wide array of applications running on leadership-class machines. 
Specifically, we study the Algebraic MultiGrid Solver (AMG), 
Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. 
The communication matrixes of each application presented in Figure~\ref{fig:amg-communication-topology}
,\ref{fig:cr-communication-topology} and \ref{fig:mg-communication-topology} are generated by
the IPM data collected from~\cite{designforwardwebpage}.
%are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.

>>>>

In this work, we select three applications from the DOE Design Forward Project. 
Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. 
We believe that the communication patterns of these applications 
are representative of a wide array of applications running on leadership-class machines. 
Specifically, we study the Algebraic MultiGrid Solver (AMG), 
Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. 
The communication matrixes of each application presented in Figure~\ref{fig:amg-communication-topology}
,\ref{fig:cr-communication-topology} and \ref{fig:mg-communication-topology} are generated by
the IPM data collected from the DOE Design Forward Project~\cite{designforwardwebpage}.
%are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.

### 2016-07-13 conclusion.tex
final pass

\section{Conclusions}
\label{sec:conclusion}
In this work, we have studied the communication behavior of three parallel applications, 
namely AMG, CrystalRouter and MultiGrid. 
Each application has distinctive communication pattern, 
that can be representative for a group of jobs in HPC workload. 
We have used a sophisticated simulation tool named CODES from Argonne National Laboratory 
to simulate the running of these three parallel applications on a torus network. 
The torus network provided by CODES has good fidelity and scalability. 
%We have analyzed the performance of each application's communication in terms of data transfer time 
%by simulating them running on torus networks with different bandwidth and dimensionality configurations. We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.

We have analyzed the intra- and inter-job interference by simulating three applications running on the torus network. 
Based on our comprehensive experiments, we got five observations. 
1) The compact allocation may not guarantee the best performance for every application. 
2) The applications dominated with nearest neighbor communications exhibit 
relatively stable performance under different allocation shapes, 
while the applications dominated with many-to-many communication exhibit 
better performance with compact allocation (e.g. 3D balanced). 
3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given. 
4) An optimal size for allocation units should be determined according to 
an application's dominant communication pattern. 
In general, a unit size should be large enough to preserve neighboring communication of the application. 
5) Inter-job interference is inevitable in non-contiguous allocation. 
However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.

We believe that the findings in this work can provide valuable guidance 
for HPC batch scheduler and resource manager to make flexible job allocations. 
Rather than using pre-defined partitions and non-contiguous placement policy, 
future HPC systems should assign each job with preferable resources based on job's communication pattern.

>>>>

\section{Conclusions}
\label{sec:conclusion}
In this work, we have studied the communication behavior of three parallel applications, 
namely AMG, CrystalRouter and MultiGrid. 
Each application has a distinctive communication pattern, 
which can be representative for a whole range of jobs commonly seen in HPC environment. 
We have used the CODES simulation toolkit  
to simulate the running of these applications on a torus network.

We have analyzed the intra- and inter-job interference by simulating three applications
running both independently and concurrently.
Based on our comprehensive experiments, we made six observations.

1) Compact allocation may not be necessary for every application.   
2) The applications dominated by nearest neighbor communications exhibit 
    relatively stable performance under different allocation shapes as long as
    the allocation exhibit some degree of locality.

3) The applications dominated with many-to-many communication exhibit 
    better performance with more compact allocations (e.g, 3D balanced).

4) A good rank-to-node mapping strategy can greatly improve 
    an application performance when a specific allocation is given.

5) An optimal size for allocation units should be determined 
    according to an application's dominant communication pattern. 
    In general, a unit size should be large enough to accommodate  
    neighboring communication in the application.

6) Inter-job interference is inevitable in non-contiguous allocation. 
    However, choosing the proper allocation unit size with communication 
    pattern awareness can help alleviate the resulting negative effects.

We believe that the findings in this work can provide valuable guidance 
for HPC batch scheduler and resource manager to make flexible job allocations. 
Rather than using pre-defined partitions or non-contiguous placement policy, 
future HPC systems should assign resources to each job based on job communication patterns.

### 2016-07-13 interference_study.tex
final pass

In this section we will study both kinds of interference 
in the context of a torus network. 
Since the applications communication patterns do not change with the scale,
\TODO{we perform our experiments with comparatively (to leadership scale) small scale,
something a 2048 rank cluster to simplify analysis.  }
%we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. 
%And in order to accommodate three applications running concurrently with different allocations, 
%we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.

@@

@@

\begin{figure*}[htp]
    \centering
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/allocation}
        \caption{ }
        \label{fig:cont_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
        \caption{ }
        \label{fig:noncont_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
        \caption{ }
        \label{fig:noncont_sub2}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
        \caption{ }
        \label{fig:noncont_sub3}
    \end{subfigure}%
    \caption{
    (a) Contiguous allocation in three different shapes. 
    ``Blue" is a 2D mesh, ``Green" is a 3D-unbalanced cube and ``Red" is a 3D-balanced cube. 
    (b)-(d) Non-contiguous allocation. Each job is represented by a specific color. 
    The nodes assigned to different jobs are interleaved, 
    the size of allocation unit are 16, 8 and 2.
    }
    \label{fig:alloc-shapes}
\end{figure*}

@@

\begin{figure*}[htp]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
        \caption{AMG}
        \label{fig:shapstudy-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
        \caption{CrystalRouter}
        \label{fig:shapstudy-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
        \caption{MultiGrid}
        \label{fig:shapstudy-mg}
    \end{subfigure}%
   \caption{
   Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 
   3D-unbalanced and 2D allocation.
   }
   \label{fig:shapestudy}
\end{figure*}

@@

AMG \TODO{remains} roughly the same when it been mapped by ``Linear" and ``Cube", 
shown in Figure~\ref{fig:diffmap-amg}. 
Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, 
both mapping strategies still preserve the locality of AMG's 3D nearest neighbor communication pattern. 
The ``Random" mapping disrupts AMG's communication pattern, 
and results in intra-job interference among the ranks. 
The performance degradation caused by using ``Random" mapping is up to 90\%.

@@

\begin{figure*}[htp]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/diffmapping/amg}
        \caption{AMG}
        \label{fig:diffmap-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/diffmapping/cr}
        \caption{CrystalRouter}
        \label{fig:diffmap-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/diffmapping/mg}
        \caption{MultiGrid}
        \label{fig:diffmap-mg}
    \end{subfigure}%
   \caption{
   Data transfer time of AMG, CrystalRouter and MultiGrid 
   on 3D-balanced allocation using  different mapping strategies.
   }
   \label{fig:diffmap}
\end{figure*}

@@

@@

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/inter-job/amg}
        \caption{AMG}
        \label{fig:interjob-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/inter-job/cr}
        \caption{CrystalRouter}
        \label{fig:interjob-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5in]{figs/inter-job/mg}
        \caption{MultiGrid}
        \label{fig:interjob-mg}
    \end{subfigure}%
   \caption{
   Inter-job interference study. 
   ``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation. 
   To study the impact of non-contiguous allocation on inter-job interference, 
   applications running concurrently with interleaved allocation of different unit sizes,
    which are 16-node, 8-node, 2-node. 
    }
   \label{fig:interjobstudy}
\end{figure*}

@@

    \item A good rank-to-node mapping strategy can greatly improve 
    an application's performance when a specific allocation is given.

    \item An optimal size for allocation units should be determined 
    according to an application's dominant communication pattern. 
    In general, a unit size should be large enough to accommodate  
    neighboring communication of the application.

>>>>

In this section we will study both kinds of interference 
in the context of a torus network. 
The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, 
such as Mira at Argonne Nation Laboratory and 
Sequoia at Lawrence Livermore National Laboratory, 
have their compute nodes connected by a 5D torus network~\cite{bgq}. 
The K computer from Japan uses the ``Tofu'' interconnected system, 
which has a 6D mesh/torus topology~\cite{tofu}. 
Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), 
has nodes connected in a 3D torus within the compute partition~\cite{titan}. 
Since the applications communication patterns do not change with the scale,
we perform our experiments with comparatively (to leadership class) modest scale,
simulating a 3D torus network with 2048 nodes ($16\times 16 \times 8$) to simplify analysis.
%we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. 
%And in order to accommodate three applications running concurrently with different allocations, 
%we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.

@@

\begin{figure}[t]
    \centering
        \includegraphics[height=2in]{figs/allocshape/allocation}
        \caption{Contiguous allocation in three different shapes. 
    ``Blue" is a 2D mesh, ``Green" is a 3D-unbalanced cube and ``Red" is a 3D-balanced cube. }
        \label{fig:cont_sub1}
\end{figure}

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/intra-job/shapestudy/amg_box}
        \caption{AMG}
        \label{fig:shapstudy-amg}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/intra-job/shapestudy/cr_box}
        \caption{CrystalRouter}
        \label{fig:shapstudy-cr}
    \end{subfigure}
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/intra-job/shapestudy/mg_box}
        \caption{MultiGrid}
        \label{fig:shapstudy-mg}
    \end{subfigure}%
   \caption{
   Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 
   3D-unbalanced and 2D allocation.
   }
   \label{fig:shapestudy}
\end{figure*}

@@

@@

AMG behavior remains roughly the same when it been mapped by ``Linear" and ``Cube", 
shown in Figure~\ref{fig:diffmap-amg}. 
Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, 
both mapping strategies still preserve the locality of AMG's 3D nearest neighbor communication pattern. 
The ``Random" mapping disrupts AMG's communication pattern, 
and results in intra-job interference among the ranks. 
The performance degradation caused by using ``Random" mapping is up to 90\%.

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/diffmapping/amg}
        \caption{AMG}
        \label{fig:diffmap-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/diffmapping/cr}
        \caption{CrystalRouter}
        \label{fig:diffmap-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/diffmapping/mg}
        \caption{MultiGrid}
        \label{fig:diffmap-mg}
    \end{subfigure}%
   \caption{
   Data transfer time of AMG, CrystalRouter and MultiGrid 
   on 3D-balanced allocation using  different mapping strategies.
   }
   \label{fig:diffmap}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
%    \begin{subfigure}[t]{0.22\textwidth}
%        \centering
%        \includegraphics[height=1.2in]{figs/allocshape/allocation}
%        \caption{ }
%        \label{fig:cont_sub1}
%    \end{subfigure}%
%    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/allocshape/unitsize/unit16}
        \caption{ }
        \label{fig:noncont_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/allocshape/unitsize/unit8}
        \caption{ }
        \label{fig:noncont_sub2}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/allocshape/unitsize/unit2}
        \caption{ }
        \label{fig:noncont_sub3}
    \end{subfigure}%
    \caption{
%    (a) Contiguous allocation in three different shapes. 
%    ``Blue" is a 2D mesh, ``Green" is a 3D-unbalanced cube and ``Red" is a 3D-balanced cube. 
    Non-contiguous allocation. Each job is represented by a specific color. 
    The nodes assigned to different jobs are interleaved, 
    the size of allocation unit are 16, 8 and 2.
    }
    \label{fig:alloc-shapes}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/inter-job/amg}
        \caption{AMG}
        \label{fig:interjob-amg}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/inter-job/cr}
        \caption{CrystalRouter}
        \label{fig:interjob-cr}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8in]{figs/inter-job/mg}
        \caption{MultiGrid}
        \label{fig:interjob-mg}
    \end{subfigure}%
   \caption{
   Inter-job interference study. 
   ``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation. 
   To study the impact of non-contiguous allocation on inter-job interference, 
   applications running concurrently with interleaved allocation of different unit sizes,
    which are 16-node, 8-node, 2-node. 
    }
   \label{fig:interjobstudy}
\end{figure*}

@@

    \item A good rank-to-node mapping strategy can greatly improve 
    an application performance when a specific allocation is given.

    \item An optimal size for allocation units should be determined 
    according to an application's dominant communication pattern. 
    In general, a unit size should be large enough to accommodate  
    neighboring communication in the application.

### 2016-07-13 intro.tex
final pass

In this work, we focus on an in-depth analysis of intra- and inter-job 
communication interference with different job placements on torus-connected HPC systems. 
\TODO{Move to Torus discussion.}
Torus-based networks are used on six of the top 10 supercomputers 
on the June 2015 Top500 lists~\cite{top500}. 
The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, 
such as Mira at Argonne Nation Laboratory and 
Sequoia at Lawrence Livermore National Laboratory, 
have their compute nodes connected by a 5D torus network~\cite{bgq}. 
The K computer from Japan uses the ``Tofu'' interconnected system, 
which has a 6D mesh/torus topology~\cite{tofu}. 
Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), 
has nodes connected in a 3D torus within the compute partition~\cite{titan}. 
Although our analyses are based on torus networks, 
the ideas conveyed in this work is applicable to networks with different topologies.

>>>>

In this work, we focus on an in-depth analysis of intra- and inter-job 
communication interference with different job placements on torus-connected HPC systems. 
Although our analyses are based on torus networks, 
the ideas conveyed in this work is applicable to networks with different topologies.

### 2016-07-13 related_work.tex
final pass

We differentiate our work from these works in the following ways. 
First, we focus on the dominant communication patterns rather than any specific application. 
We believe this can provide a guideline for other research work. 
Secondly, we explore \TODO{the} intra- and inter-job interference between concurrently running jobs, 
while similar work such as~\cite{abhinav-sc13} focuses 
on one single application's performance degradation due to network contention. 
Finally, we analyze the impact of different placement strategies to job's communication behaviors, 
identifying preferred placement strategies for 
each application with a specific dominant communication pattern. 
Based on our study, we claim that future batch schedulers should take 
job communication patterns into consideration for placement decision making.

>>>>

We differentiate our work from these works in the following ways. 
First, we focus on the dominant communication patterns rather than any specific application. 
We believe this can provide a guideline for other research work. 
Secondly, we explore both intra- and inter-job interference between concurrently running jobs, 
while similar work such as~\cite{abhinav-sc13} focuses 
on one single application's performance degradation due to network contention. 
Finally, we analyze the impact of different placement strategies to job's communication behaviors, 
identifying preferred placement strategies for 
each application with a specific dominant communication pattern. 
Based on our study, we claim that future batch schedulers should take 
job communication patterns into consideration for placement decision making.
